Tutorials: A Brief Overview

### 2023-04-25 ListofReferences.qmd
css updates

GLM @sec-glm
Non-independence @Non-independence

## GLM (Generalized Linear Model): 
A flexible generalization of ordinary linear regression that allows for response variables that have error distribution models other than a normal distribution.{#sec-glm}

@@

3.	Sample average of differences: The mean of the differences between paired observations, typically used in paired-sample designs to assess the average change between two time points or conditions.

4.	Fixed effects: Effects in a statistical model that are constant across all observations, representing the average relationship between predictors and the outcome variable.

5.	Random effects: Effects in a statistical model that vary across different levels of a grouping factor (e.g., subjects, clusters), capturing the variability within each level.

6.	Hierarchical data structures: Data structures where observations are nested within higher-order units (e.g., students within schools, employees within companies), resulting in dependency among observations.

7.	Within-subject correlations: Correlations between repeated measurements on the same subject or unit, often present in longitudinal or clustered data.

8.	Model fit indices: Quantitative measures used to assess how well a statistical model fits the data, such as R-squared, AIC, or BIC.

9.	Nonparametric: Statistical methods that do not rely on assumptions about the underlying distribution of the data.

10.	Resampling permutation method: A nonparametric method for hypothesis testing that involves randomly rearranging the observed data to create new samples and assess the probability of the observed test statistic under the null hypothesis.

11.	Clustered data: Data where observations are grouped or clustered within higher-order units, resulting in dependency among observations within each cluster.

12.	Population-averaged effects: Effects in a statistical model that represent the average relationship between predictors and the outcome variable across the entire population, rather than for specific levels of a grouping factor.

13.	Identity link: A link function in GLMs that relates the linear predictor to the expected value of the outcome variable without any transformation (E[Y] = Xβ).

14.	Logit link: A link function in GLMs used for binary outcomes, relating the linear predictor to the log-odds of the outcome variable (log(E[Y]/(1-E[Y])) = Xβ).

15.	Log link: A link function in GLMs used for count outcomes, relating the linear predictor to the natural logarithm of the expected value of the outcome variable (log(E[Y]) = Xβ).

16.	Logistic link: Synonymous with the logit link.

17.	Binomial distribution: A probability distribution that describes the number of successes in a fixed number of Bernoulli trials with the same probability of success.

18.	Exchangeable correlation structure: A working correlation structure in GEE models, assuming constant correlation between any pair of repeated measurements within subjects.

19.	Autoregressive correlation structure: A working correlation structure in GEE models, assuming correlation between repeated measurements decreases as the time lag between measurements increases.

20.	Unstructured correlation structure: A working correlation structure in GEE models, allowing for different correlations between each pair of repeated measurements within subjects.

21.	Sandwich estimator: A method for computing robust standard errors in GEE models, providing valid inference even if the working correlation structure is misspecified.

22.	Quasi-likelihood under the Independence model Criterion (QIC): A model selection criterion for GEE models, analogous to AIC, used to compare models with different working correlation structures.

23.	Odds ratios: A measure of association for binary outcomes, representing the ratio of the odds of an event occurring in one group to the odds in another group.

24.	Rate ratios: A measure of association for count outcomes, representing the ratio of the rates of an event occurring in one group to the rates in another group.

25.	Maximum likelihood estimation: A statistical method for estimating model parameters by finding the values that maximize the likelihood of the observed data given the model.

26.	Unbalanced data: Data where the number of observations or the number of measurements per subject or group is unequal across subjects or groups.

27.	Unequally spaced data: Data where the time intervals between repeated measurements or the spacing of observations are not equal.

28.	Maximum likelihood (ML): A method of estimation that involves finding the parameter values that maximize the likelihood of the observed data given a statistical model.

29.	Restricted maximum likelihood (REML): A modification of maximum likelihood estimation that is less biased for estimating variance components in mixed-effects models, particularly when the sample size is small.

30.	Penalized quasi-likelihood (PQL): An approximate method for fitting generalized linear mixed models, particularly for non-normal outcomes, which involves a penalized likelihood function and iteratively weighted least squares.

31.	Akaike Information Criterion (AIC): A model selection criterion that balances goodness-of-fit and model complexity, calculated as AIC = -2 * log-likelihood + 2 * k, where k is the number of parameters. Lower AIC values indicate better model fit.

32.	Bayesian Information Criterion (BIC): A model selection criterion that balances goodness-of-fit and model complexity, calculated as BIC = -2 * log-likelihood + k * log(n), where k is the number of parameters and n is the sample size. Lower BIC values indicate better model fit.

>>>>

## Generalized Linear Model (GLM): 
A flexible generalization of ordinary linear regression that allows for response variables that have error distribution models other than a normal distribution.{#sec-glm}

@@

## Sample average of differences: 
The mean of the differences between paired observations, typically used in paired-sample designs to assess the average change between two time points or conditions.

## Fixed effects: 
Effects in a statistical model that are constant across all observations, representing the average relationship between predictors and the outcome variable.

## Random effects: 
Effects in a statistical model that vary across different levels of a grouping factor (e.g., subjects, clusters), capturing the variability within each level.

## Hierarchical data structures: 
Data structures where observations are nested within higher-order units (e.g., students within schools, employees within companies), resulting in dependency among observations.

## Within-subject correlations: 
Correlations between repeated measurements on the same subject or unit, often present in longitudinal or clustered data.

## Model fit indices: 
Quantitative measures used to assess how well a statistical model fits the data, such as R-squared, AIC, or BIC.

## Nonparametric: 
Statistical methods that do not rely on assumptions about the underlying distribution of the data.

## Resampling permutation method: 
A nonparametric method for hypothesis testing that involves randomly rearranging the observed data to create new samples and assess the probability of the observed test statistic under the null hypothesis.

## Clustered data: 
Data where observations are grouped or clustered within higher-order units, resulting in dependency among observations within each cluster.

## Population-averaged effects: 
Effects in a statistical model that represent the average relationship between predictors and the outcome variable across the entire population, rather than for specific levels of a grouping factor.

## Identity link: 
A link function in GLMs that relates the linear predictor to the expected value of the outcome variable without any transformation (E[Y] = Xβ).

## Logit link: 
A link function in GLMs used for binary outcomes, relating the linear predictor to the log-odds of the outcome variable (log(E[Y]/(1-E[Y])) = Xβ).

## Log link: 
A link function in GLMs used for count outcomes, relating the linear predictor to the natural logarithm of the expected value of the outcome variable (log(E[Y]) = Xβ).

## Logistic link: 
Synonymous with the logit link.

## Binomial distribution: 
A probability distribution that describes the number of successes in a fixed number of Bernoulli trials with the same probability of success.

## Exchangeable correlation structure: 
A working correlation structure in GEE models, assuming constant correlation between any pair of repeated measurements within subjects.

## Autoregressive correlation structure: 
A working correlation structure in GEE models, assuming correlation between repeated measurements decreases as the time lag between measurements increases.

## Unstructured correlation structure: 
A working correlation structure in GEE models, allowing for different correlations between each pair of repeated measurements within subjects.

## Sandwich estimator: 
A method for computing robust standard errors in GEE models, providing valid inference even if the working correlation structure is misspecified.

## Quasi-likelihood under the Independence model Criterion (QIC): 
A model selection criterion for GEE models, analogous to AIC, used to compare models with different working correlation structures.

## Odds ratios: 
A measure of association for binary outcomes, representing the ratio of the odds of an event occurring in one group to the odds in another group.

## Rate ratios: 
A measure of association for count outcomes, representing the ratio of the rates of an event occurring in one group to the rates in another group.

## Maximum likelihood estimation: 
A statistical method for estimating model parameters by finding the values that maximize the likelihood of the observed data given the model.

## Unbalanced data: 
Data where the number of observations or the number of measurements per subject or group is unequal across subjects or groups.

## Unequally spaced data: 
Data where the time intervals between repeated measurements or the spacing of observations are not equal.

## Maximum likelihood (ML): 
A method of estimation that involves finding the parameter values that maximize the likelihood of the observed data given a statistical model.

## Restricted maximum likelihood (REML): 
A modification of maximum likelihood estimation that is less biased for estimating variance components in mixed-effects models, particularly when the sample size is small.

## Penalized quasi-likelihood (PQL): 
An approximate method for fitting generalized linear mixed models, particularly for non-normal outcomes, which involves a penalized likelihood function and iteratively weighted least squares.

## Akaike Information Criterion (AIC): 
A model selection criterion that balances goodness-of-fit and model complexity, calculated as AIC = -2 * log-likelihood + 2 * k, where k is the number of parameters. Lower AIC values indicate better model fit.

## Bayesian Information Criterion (BIC): 
A model selection criterion that balances goodness-of-fit and model complexity, calculated as BIC = -2 * log-likelihood + k * log(n), where k is the number of parameters and n is the sample size. Lower BIC values indicate better model fit.

### 2023-04-25 explanations.qmd
css updates

---
title: "Explanations: A Brief Overview"
---

The "Explanation" section focuses on providing context, background information, and a deeper understanding of the concepts, methodologies, or processes related to a specific analyses or topic. Unlike tutorials and how-to guides, which are goal-oriented and problem-oriented respectively, explanations aim to clarify and enhance the user's knowledge. The explanation section aims to create informative, engaging, and accessible content that helps users build a solid foundation of understanding and knowledge about the subject matter.

@@

1. Be informative: Explanations should offer valuable insights and information, helping users develop a deeper understanding of the topic. This includes discussing the rationale behind certain choices, exploring alternative approaches, and explaining the advantages and disadvantages of various options.

2. Address "why" questions: Explanations should aim to answer the "why" questions users may have, delving into the reasoning, motivations, or principles that underlie a specific concept or approach.

3. Be clear and concise: While explanations may go into greater depth than tutorials or how-to guides, they should still be presented in a clear and concise manner. This means using plain language, avoiding jargon, and providing examples to help users grasp the concepts more easily.

4. Be well-structured: Explanations should be organized in a logical and coherent manner, making it easy for users to follow the flow of information and understand the relationships between different concepts.

5. Support other documentation types: Explanations should complement and support other types of documentation, such as tutorials and how-to guides, by providing the necessary context and background information. This may involve cross-referencing other resources or linking to related documentation.

6. Engage the audience: Explanations should be engaging and interesting, capturing the user's attention and encouraging them to explore the topic further. This may involve using analogies, storytelling, or real-world examples to illustrate the concepts and make them more relatable.

For additional information/resources, see [https://documentation.divio.com/]

>>>>

---
title: "Explanations: A Brief Overview"
number-sections: true
format:
 html:
   css: styles.css
   theme: cosmo
   fontsize: 1.1em
   linestretch: 1.5
---

## What are Explanations?

::: {style="text-align: justify; font-size: 1.2em; line-stretch: 1.o; font-family: Georgia, serif;"}
[**The primary goal**]{style="color: goldenrod; font-size: 1.2em; font-family: Verdana, sans-serif;"} 
of the explanation section is to provide context, background information, and a deeper understanding of the concepts, methodologies, or processes related to a specific analysis or topic. Unlike tutorials and how-to guides, which are goal-oriented and problem-oriented respectively, explanations aim to clarify and enhance the user's knowledge. The explanation section aims to create informative, engaging, and accessible content that helps users build a solid foundation of understanding and knowledge about the subject matter.
:::

::: {style="text-align: justify; line-height: normal; color: black; border: solid black; background-color: antiquewhite; padding: 70px; font-family: Georgia, serif;"}

@@

1. **Be informative**: Explanations should offer valuable insights and information, helping users develop a deeper understanding of the topic. This includes discussing the rationale behind certain choices, exploring alternative approaches, and explaining the advantages and disadvantages of various options.

2. **Address "why" questions**: Explanations should aim to answer the "why" questions users may have, delving into the reasoning, motivations, or principles that underlie a specific concept or approach.

3. **Be clear and concise**: While explanations may go into greater depth than tutorials or how-to guides, they should still be presented in a clear and concise manner. This means using plain language, avoiding jargon, and providing examples to help users grasp the concepts more easily.

4. **Be well-structured**: Explanations should be organized in a logical and coherent manner, making it easy for users to follow the flow of information and understand the relationships between different concepts.

5. **Support other documentation types**: Explanations should complement and support other types of documentation, such as tutorials and how-to guides, by providing the necessary context and background information. This may involve cross-referencing other resources or linking to related documentation.

6. **Engage the audience**: Explanations should be engaging and interesting, capturing the user's attention and encouraging them to explore the topic further. This may involve using analogies, storytelling, or real-world examples to illustrate the concepts and make them more relatable.

::: {style="text-align: center"}
For additional information, see [https://documentation.divio.com/]
:::

:::

### 2023-04-25 howto.qmd
css updates

---
title: "How-To's: A Brief Overview"
---

[Section in progress...]

>>>>

---
title: "How-To's: A Brief Overview"
number-sections: true
format:
 html:
   css: styles.css
   theme: cosmo
   fontsize: 1.1em
   linestretch: 1.5
---

## What are How-To's?
::: {style="text-align: justify; font-size: 1.2em; line-stretch: 1.o; font-family: Georgia, serif;"}
[**The primary goal**]{style="color: goldenrod; font-size: 1.2em; font-family: Verdana, sans-serif;"} 
of the "how-to" section is on providing practical solutions to specific methods or analytic problems users might encounter. Unlike tutorials, which aim to teach a concept, how-to guides concentrate on the efficient execution of specific analyses without going into in-depth background explanations. By following these principles, the how-to section aims to create practical, efficient, and user-friendly guides that help users quickly identify solutions and accomplish their goals.
:::

::: {style="text-align: justify; line-height: normal; color: black; border: solid black; background-color: antiquewhite; padding: 70px; font-family: Georgia, serif;"}

The how-to section should:

1. **Be problem-oriented**: How-to guides should address specific problems or analyses users may face, offering clear and concise solutions. Each guide should focus on a single issue, making it easy for users to find and apply the relevant information.

2. **Offer step-by-step instructions**: Similar to tutorials, how-to guides should provide a clear sequence of steps that users can follow to solve their problem or complete their analysis. Each step of the instructions should be explained in a straightforward manner.

3. **Be concise**: How-to guides should be to the point, offering just enough information to help users accomplish their goal. This means avoiding unnecessary explanations or background information, as users are typically looking for a quick solution.

4. **Be actionable**: The content in the how-to section should be practical, with a focus on actionable advice that users can apply immediately. This may involve providing code snippets, configuration settings, or other specific instructions that users can implement in their projects.

5. **Include examples**: How-to guides should include relevant examples that demonstrate the solution in action, allowing users to see how the instructions can be applied in a real-world context.

6. **Cross-reference other resources**: How-to guides should link to other relevant documentation, such as explanations or reference material, for users who may require additional background information or a deeper understanding of the topic.

::: {style="text-align: center"}
For additional information, see [https://documentation.divio.com/]
:::

:::

### 2023-04-25 index.qmd
css updates

---
#title: "Temporary Landing Page for Longitudinal Project"
#format:
#  html:
#    css: styles.css
#    theme: cosmo
#    fontsize: 1.1em
    #linestretch: 1.7
---

<div class="text-3xl text-center">
This is the temporary landing page for the Longitudinal Analysis Project. <br> The site is currently under construction and used for testing purposes only. <br> Contents may be inaccurate.
</div>

>>>>

---
number-sections: true
#title: "Longitudinal Analysis"
format:
  html:
    css: styles.css
    theme: cosmo
    fontsize: 1.3em
    linestretch: 1.7
---

::: {style="text-align: justify; text-align: center; line-height: normal; color: white; border: solid black; background-color: darkgoldenrod; padding: 70px; font-family: Georgia, serif;"}

This is the temporary landing page for the Longitudinal Analysis Project. <br> The site is currently under construction and used for testing purposes only. <br> Contents may be inaccurate.

### 2023-04-25 reference.qmd
css updates

---
title: "Reference Material: A Brief Overview"
---

[Section in progress...]

>>>>

---
title: "References: A Brief Overview"
number-sections: true
format:
 html:
   css: styles.css
   theme: cosmo
   fontsize: 1.1em
   linestretch: 1.5
---

## What are References?

::: {style="text-align: justify; font-size: 1.2em; line-stretch: 1.o; font-family: Georgia, serif;"}
[**The primary goal**]{style="color: goldenrod; font-size: 1.2em; font-family: Verdana, sans-serif;"} 
of the "Reference" section is on providing accurate, concise, and structured information about specific components, functions, or features of a method, analysis, or framework. Unlike tutorials, how-to guides, and explanations, which aim to teach, solve problems, and provide context, respectively, reference materials serve as a quick lookup resource for users to find the exact information they need. By following these principles, the reference section aims to create a reliable, user-friendly, and easily searchable resource that provides users with the specific information they need to effectively use and understand the method, analysis, or framework.
:::

::: {style="text-align: justify; line-height: normal; color: black; border: solid black; background-color: antiquewhite; padding: 70px; font-family: Georgia, serif;"}

The reference section should:

1. **Be comprehensive**: Reference materials should cover all relevant aspects of the method, analysis, or framework, including components, functions, features, and configurations. The goal is to provide users with a complete and authoritative source of information.

2. **Be accurate**: The information in the reference section should be precise and up-to-date, ensuring that users can rely on it for their projects. This may involve regular updates and checks to maintain accuracy.

3. **Be concise**: Reference materials should be brief and to the point, providing just enough information for users to understand and use the specific analysis or method. This means avoiding unnecessary explanations, examples, or background information.

4. **Follow a consistent format**: The reference section should use a consistent format and structure, making it easy for users to locate and access the information they need. This may involve using templates, standardized headings, or other organizational tools.

5. **Be easily searchable**: Reference materials should be designed with searchability in mind, enabling users to quickly find the information they need. This may involve using clear and descriptive titles, headings, and keywords, as well as implementing search functionality or indexing.

6. **Cross-reference other resources**: The reference section should link to other relevant documentation, such as tutorials, how-to guides, or explanations, for users who may require additional information or a deeper understanding of the topic.

::: {style="text-align: center"}
For additional information, see [https://documentation.divio.com/]
:::

:::

### 2023-04-25 tutorials.qmd
css updates

---
title: "Tutorials: A Brief Overview"
---

[Section in progress...]

>>>>

---
title: "Tutorials: A Brief Overview"
number-sections: true
format:
 html:
   css: styles.css
   theme: cosmo
   fontsize: 1.1em
   linestretch: 1.5
---

## What are Tutorials?

::: {style="text-align: justify; font-size: 1.2em; line-stretch: 1.o; font-family: Georgia, serif;"}
[**The primary goal**]{style="color: goldenrod; font-size: 1.2em; font-family: Verdana, sans-serif;"} 
of the tutorial section is to help users learn and understand a specific topic or method by providing an introductory and hands-on, step-by-step guide. This section aims to create engaging, accessible, and effective entry learning experiences that enable users to develop their skills and understanding of the subject matter. 
:::

::: {style="text-align: justify; line-height: normal; color: black; border: solid black; background-color: antiquewhite; padding: 70px; font-family: Georgia, serif;"}

The tutorial section should:

1. **Be goal-oriented**: Tutorials should focus on a specific task the user wants to accomplish. By concentrating on a particular outcome, the tutorial becomes more engaging and relevant to the user.

2. **Provide clear steps**: Tutorials should offer step-by-step instructions, guiding users through the process of achieving their goal. Each step should be clearly explained and easy to follow, making sure users can successfully complete the tutorial.

3. **Use examples**: Tutorials should include real-world examples and use cases that are relatable and meaningful to the user. This helps users better understand the concepts being taught and how they can apply them in their own projects.

4. **Be practical**: Tutorials should emphasize hands-on learning, encouraging users to actively engage with the content and practice the skills they are learning. This may involve providing code samples, interactive exercises, or other activities that help users develop their understanding.

5. **Encourage exploration**: Tutorials should encourage users to explore and experiment with the concepts they are learning. This may involve providing opportunities for users to adapt the tutorial to their own needs or offering suggestions for further experimentation.

::: {style="text-align: center"}
For additional information, see [https://documentation.divio.com/]
:::

:::
